---
title: "Assignment 4"
author: "Soha Arian"
format:
  html:
    toc: true
    theme: cosmo
    css: styles.css
execute:
  echo: true
  warning: false
  message: false
---

# Assignment 4

## Text Analytics using quanteda

## 1. Read tutorial of quanteda

The quanteda tutorial provides the workflow used in this assignment: corpus, tokens, and document-feature matrix.

## 2a. Use sample programs to analyze US presidential inaugural speeches

This section uses the built-in inaugural speech corpus in quanteda.

```{r}
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textstats)
library(quanteda.textplots)
library(readtext)
library(ggplot2)

theme_set(
  theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(color = "#d63384", face = "bold"),
      plot.subtitle = element_text(color = "#d63384"),
      axis.title = element_text(color = "#d63384"),
      axis.text = element_text(color = "#d63384"),
      legend.title = element_text(color = "#d63384"),
      legend.text = element_text(color = "#d63384"),
      strip.text = element_text(color = "#d63384", face = "bold")
    )
)

packageVersion("quanteda")
```

```{r}
corp_inaug <- data_corpus_inaugural
summary(corp_inaug, n = 5)
head(docvars(corp_inaug))
```

```{r}
toks_inaug <- tokens(
  corp_inaug,
  remove_punct = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE
)

toks_inaug <- tokens_remove(toks_inaug, pattern = stopwords("en"))

dfmat_inaug <- dfm(toks_inaug)
dfmat_inaug
```

### Similarities and differences among presidents

```{r}
dfmat_pres <- dfm_group(dfmat_inaug, groups = President)
topfeatures(dfmat_pres, 20)
```

```{r}
pres_sim <- textstat_simil(dfmat_pres, method = "cosine")
as.matrix(pres_sim)
```

### Similarities and differences over time

```{r}
era <- ifelse(docvars(corp_inaug, "Year") < 1900, "Before1900", "1900plus")
dfmat_era <- dfm_group(dfmat_inaug, groups = era)
topfeatures(dfmat_era, 20)
```

```{r}
key_era <- textstat_keyness(dfmat_era, target = "1900plus")
head(key_era, 20)
```

```{r}
textplot_keyness(
  key_era,
  color = c("#d63384", "#f3bfd7"),
  labelcolor = "#d63384"
) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(color = "#d63384", face = "bold"),
    axis.text = element_text(color = "#d63384"),
    axis.title = element_text(color = "#d63384"),
    legend.text = element_text(color = "#d63384"),
    legend.title = element_text(color = "#d63384")
  )
```

## 2b. Use sample programs to analyze US Department of Defense reports

This section reads the DOD reports from the USDOD folder.

```{r}
dod_texts <- readtext("USDOD/*.pdf")
dod_texts
```

```{r}
corp_dod <- corpus(dod_texts)
summary(corp_dod)
```

```{r}
toks_dod <- tokens(
  corp_dod,
  remove_punct = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE
)

toks_dod <- tokens_remove(toks_dod, pattern = stopwords("en"))

dfmat_dod <- dfm(toks_dod)
dfmat_dod_trim <- dfm_trim(dfmat_dod, min_termfreq = 5)

dfmat_dod_trim
```

```{r}
topfeatures(dfmat_dod_trim, 30)
```

```{r}
dod_sim <- textstat_simil(dfmat_dod_trim, method = "cosine")
as.matrix(dod_sim)
```

## 3. What is Wordfish?

Wordfish is an unsupervised one-dimensional text scaling method that estimates document positions from observed word frequencies.

## 3a. Run Wordfish on DOD reports

```{r}
wf_dod <- textmodel_wordfish(dfmat_dod_trim, dir = c(1, ndoc(dfmat_dod_trim)))
summary(wf_dod)
```

```{r}
wf_plot_data <- data.frame(
  document = docnames(dfmat_dod_trim),
  wordfish_position = as.numeric(wf_dod$theta)
)

ggplot(wf_plot_data, aes(x = wordfish_position, y = reorder(document, wordfish_position))) +
  geom_point(size = 3, color = "#d63384") +
  labs(
    title = "Wordfish Positions for DOD Reports",
    x = "Wordfish Position",
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(color = "#d63384", face = "bold"),
    axis.text = element_text(color = "#d63384"),
    axis.title = element_text(color = "#d63384")
  )
```

```{r}
dod_wordfish_positions <- data.frame(
  document = docnames(dfmat_dod_trim),
  wordfish_position = as.numeric(wf_dod$theta)
)

knitr::kable(dod_wordfish_positions, digits = 3, format = "html")
```

## 4. How to compare positions

Positions can be compared by examining the estimated document scores from Wordfish and comparing them to another scaling method such as correspondence analysis.

```{r}
ca_dod <- textmodel_ca(dfmat_dod_trim)
```

```{r}
dod_positions_compare <- data.frame(
  document = docnames(dfmat_dod_trim),
  wordfish = as.numeric(wf_dod$theta),
  ca_dim1 = as.numeric(coef(ca_dod, doc_dim = 1)$coef_document)
)

knitr::kable(dod_positions_compare, digits = 3, format = "html")
```

```{r}
ggplot(dod_positions_compare, aes(x = wordfish, y = ca_dim1, label = document)) +
  geom_point(size = 3, color = "#d63384") +
  geom_text(size = 3, vjust = -0.5, check_overlap = TRUE, color = "#d63384") +
  labs(
    title = "Comparing Document Positions",
    x = "Wordfish Position",
    y = "Correspondence Analysis Dimension 1"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(color = "#d63384", face = "bold"),
    axis.text = element_text(color = "#d63384"),
    axis.title = element_text(color = "#d63384")
  )
```

```{r}
cor(dod_positions_compare$wordfish, dod_positions_compare$ca_dim1)
```

**AI Disclaimer:** *I used ChatGPT as a tutor to clarify instructions, troubleshoot errors, and better understand mistakes encountered during the process. It also assisted with wording and formatting; however, the final submission reflects my own work and judgment.*

```{=html}
<style>

/* Make all Assignment 4 regular text white */
main.content,
main.content p,
main.content li,
main.content span,
main.content div,
main.content label {
  color: #ffffff !important;
  -webkit-text-fill-color: #ffffff !important;
}

/* Make all Assignment 4 headings white */
main.content h1,
main.content h2,
main.content h3,
main.content h4,
main.content h5,
main.content h6 {
  color: #ffffff !important;
  -webkit-text-fill-color: #ffffff !important;
}

/* Make all R output text white */
main.content .cell-output,
main.content .cell-output-stdout,
main.content .cell-output-display,
main.content .cell-output-stderr,
main.content .cell-output pre,
main.content .cell-output code,
main.content .cell-output-stdout pre,
main.content .cell-output-stdout code,
main.content pre:not(.sourceCode),
main.content pre:not(.sourceCode) code {
  color: #ffffff !important;
  -webkit-text-fill-color: #ffffff !important;
}

/* Make printed tables and data output white */
main.content .cell-output table,
main.content .cell-output th,
main.content .cell-output td,
main.content table,
main.content th,
main.content td {
  color: #ffffff !important;
  -webkit-text-fill-color: #ffffff !important;
}

/* Keep R source code readable */
main.content pre.sourceCode,
main.content pre.sourceCode code,
main.content .sourceCode {
  color: #f2f2ed !important;
  -webkit-text-fill-color: #f2f2ed !important;
}

/* Keep links bright green */
main.content a {
  color: #dfff54 !important;
  -webkit-text-fill-color: #dfff54 !important;
}

</style>
```

```{=html}
<style>

/* Make graph titles and captions white, bold, and uppercase */
main.content .figure-caption,
main.content figcaption,
main.content .cell-output-display {
  color: #ffffff !important;
  font-weight: 800 !important;
  text-transform: uppercase !important;
}

/* Style text inside SVG graphs */
main.content .cell-output-display svg text {
  fill: #ffffff !important;
  color: #ffffff !important;
  font-weight: 800 !important;
  text-transform: uppercase !important;
  font-family: Arial, Helvetica, sans-serif !important;
}

/* Make graph titles larger */
main.content .cell-output-display svg .title {
  font-size: 18px !important;
  font-weight: 800 !important;
}

/* Make graph axis titles bold */
main.content .cell-output-display svg .axis-title {
  font-weight: 800 !important;
  text-transform: uppercase !important;
}

/* Make graph legend text bold */
main.content .cell-output-display svg .legend {
  font-weight: 800 !important;
  text-transform: uppercase !important;
}

</style>
```

library(ggplot2) library(ggrepel) library(dplyr) library(stringr)

# Clean and capitalize document names

plot_data \<- plot_data \|\> mutate( document_label = document \|\> basename() \|\> str_remove("\\.pdf\$") \|\> str_replace_all("\_", " ") \|\> str_squish() \|\> str_to_upper() )

# Create a cleaner graph

ggplot( plot_data, aes( x = wordfish_position, y = ca_dimension_1 ) ) + geom_point( color = "#d11973", size = 4, alpha = 0.9 ) + geom_text_repel( aes(label = document_label), color = "#d11973", size = 4, fontface = "bold", box.padding = 0.7, point.padding = 0.5, min.segment.length = 0, max.overlaps = Inf, segment.color = "#8c8c8c", segment.linewidth = 0.35 ) + labs( title = "COMPARING DOCUMENT POSITIONS", subtitle = "WORDFISH AND CORRESPONDENCE ANALYSIS RESULTS", x = "WORDFISH POSITION", y = "CORRESPONDENCE ANALYSIS — DIMENSION 1" ) + theme_minimal(base_size = 14) + theme( text = element_text( family = "Arial", face = "bold" ),

```         
plot.title = element_text(
  color = "#d11973",
  size = 24,
  face = "bold",
  hjust = 0.5,
  margin = margin(bottom = 8)
),

plot.subtitle = element_text(
  color = "#555555",
  size = 12,
  face = "bold",
  hjust = 0.5,
  margin = margin(bottom = 20)
),

axis.title = element_text(
  color = "#d11973",
  size = 15,
  face = "bold"
),

axis.text = element_text(
  color = "#333333",
  size = 11,
  face = "bold"
),

panel.grid.major = element_line(
  color = "#dddddd",
  linewidth = 0.6
),

panel.grid.minor = element_blank(),

plot.margin = margin(
  25,
  55,
  25,
  35
)
```

)
