---
title: "Assignment 1 – Brush up R and Quarto"
format: html
---

# Exploratory Data Analysis

I reviewed Chapters 3 to 7 in R4DS, Section 3.5 in Data Programming. Using this information, I will run an exploratory data analysis with R using the TEDS2016 dataset.

## Import the dataset

This code downloads the TEDS2016 file in Stata format and loads it into R.

```{r}
library(haven)
library(ggplot2)

TEDS_2016 <- haven::read_dta("https://github.com/datageneration/home/blob/master/DataProgramming/data/TEDS_2016.dta?raw=true")
```

## Problems I See In The Dataset

I noticed that some variables are stored as coded numbers instead of clear words, which makes them harder to interpret until labels are added. I also found missing values in multiple columns, which can change results if they are not handled. Finally, some answers are non substantive, such as no response.

## Dealing With Missing Values

Missing values can be handled by removing rows that have missing values for the variables being analyzed or by replacing missing numeric values with a reasonable value such as the mean. For this assignment, I remove rows where Tondu is missing because Tondu is required for the next steps. I also replace missing age values with the average age so age can still be used.

### Check For Missing Values

```{r}
colSums(is.na(TEDS_2016))
```

### Remove Incomplete Rows

```{r}
TEDS_clean <- TEDS_2016[, c("Tondu","female","DPP","age","income","edu","Taiwanese","Econ_worse")]
TEDS_clean <- na.omit(TEDS_clean)
```

## Relationship Between Tondu And Other Variables

To explore Tondu with categorical variables such as female, DPP, Taiwanese, and Econ_worse, I use cross tabulations and proportions to compare group patterns. To explore Tondu with numeric or ordered variables such as age, income, and education, I compare averages by Tondu category. These methods provide a clear first look at differences across groups.

### Use Facets From R For Data Science (Section 3.5).

#### Tondu by Female

```{r}
ggplot(TEDS_clean, aes(x = Tondu)) +
  geom_bar(fill = "deeppink") +
  facet_wrap(~ female)
```

#### Tondu by Female and DPP

```{r}
ggplot(TEDS_clean, aes(x = Tondu)) +
  geom_bar(fill = "deeppink") +
  facet_grid(female ~ DPP)
```

## Looking At Votetsai

The votetsai variable shows whether a respondent voted for Tsai Ing wen. I first check the overall counts for votetsai, then compare it with Tondu using a cross tabulation to see how voting relates to Tondu attitudes.

### Find the correct votesai column name

```{r}
names(TEDS_2016)
```

## Frequency Table And Bar Chart For Tondu

### Assign labels

```{r}
TEDS_2016$Tondu <- as.numeric(
  TEDS_2016$Tondu,
  labels = c(
    "Unification now",
    "Status quo, unif. in future",
    "Status quo, decide later",
    "Status quo forever",
    "Status quo, indep. in future",
    "Independence now",
    "No response"
  )
)
```

```{r}
table(TEDS_2016$Tondu)
```

```{r}
ggplot(TEDS_2016, aes(x = Tondu)) +
  geom_bar(fill = "deeppink")
```

```{r}
library(haven)
library(ggplot2)

TEDS_2016 <- haven::read_dta("https://github.com/datageneration/home/blob/master/DataProgramming/data/TEDS_2016.dta?raw=true")

TEDS_clean <- TEDS_2016[, c("Tondu","female","DPP","age","income","edu","Taiwanese","Econ_worse","votetsai")]
TEDS_clean <- na.omit(TEDS_clean)

ggplot(TEDS_clean, aes(x = Tondu)) +
  geom_bar(fill = "deeppink") +
  facet_wrap(~ votetsai)
```