---
title: "Assignment 3: ML Exercise"
author: "Soha Arian"
format: html
execute:
  echo: true
---

## 1. Run Lab_ML01.R

The assignment requires running `Lab_ML01.R` first. This serves as preparation for the exercise because it introduces the lab workflow and helps establish the regression and plotting format used in the assignment.

```{r}
# ============================================================================
# Workshop: Machine Learning with Survey Data (TEDS 2016)
# EPPS 6323 Knowledge Mining
# Karl Ho, University of Texas at Dallas
# ============================================================================

# ----------------------------------------------------------------------------
# 0. Setup: Packages and Data
# ----------------------------------------------------------------------------

library(haven)        # Read Stata files
library(tidyverse)    # Data wrangling and visualization
library(GGally)       # Pairs plots
library(cluster)      # Clustering
library(factoextra)   # Visualize clusters and PCA
library(rpart)        # Decision trees
library(rpart.plot)   # Plot decision trees
library(randomForest) # Random forest
library(caret)        # Model training and evaluation
library(e1071)        # SVM and Naive Bayes

# Load data
TEDS_2016 <- read_stata("https://github.com/datageneration/home/blob/master/DataProgramming/data/TEDS_2016.dta?raw=true")


# ============================================================================
# PART I: EXPLORATORY DATA ANALYSIS
# ============================================================================

# ----------------------------------------------------------------------------
# Q1. Data overview
# ----------------------------------------------------------------------------

dim(TEDS_2016)
names(TEDS_2016)
summary(TEDS_2016)

# ----------------------------------------------------------------------------
# Q2. Missing data assessment
# ----------------------------------------------------------------------------

TEDS_2016 %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), 
               names_to = "variable", 
               values_to = "n_missing") %>%
  filter(n_missing > 0) %>%
  arrange(desc(n_missing))

# Visualize missingness
TEDS_2016 %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), 
               names_to = "variable", 
               values_to = "n_missing") %>%
  filter(n_missing > 0) %>%
  ggplot(aes(x = reorder(variable, n_missing), y = n_missing)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Missing Values by Variable",
       x = NULL, y = "Count of Missing Values") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q3. Variable recoding
# ----------------------------------------------------------------------------

teds <- TEDS_2016 %>%
  mutate(
    vote = factor(votetsai, 
                  levels = c(0, 1), 
                  labels = c("Other", "Tsai")),
    gender = factor(female, 
                    levels = c(0, 1), 
                    labels = c("Male", "Female")),
    Tondu = as.factor(Tondu),
    Party = as.factor(Party)
  ) %>%
  dplyr::select(vote, gender, age, edu, income, 
         Taiwanese, Econ_worse, Tondu, Party, DPP) %>%
  drop_na()

glimpse(teds)

# ----------------------------------------------------------------------------
# Q4. Vote choice distribution
# ----------------------------------------------------------------------------

ggplot(teds, aes(x = vote, fill = vote)) +
  geom_bar() +
  scale_fill_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  labs(title = "Vote Choice Distribution",
       subtitle = "TEDS 2016: Tsai Ing-wen vs. Other Candidates",
       x = "Vote Choice", y = "Count") +
  theme_minimal(base_size = 18) +
  theme(legend.position = "none")

# ----------------------------------------------------------------------------
# Q5. Age distribution
# ----------------------------------------------------------------------------

ggplot(teds, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  labs(title = "Age Distribution of Respondents",
       x = "Age", y = "Count") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q6. Vote by gender
# ----------------------------------------------------------------------------

ggplot(teds, aes(x = gender, fill = vote)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  labs(title = "Vote Choice by Gender",
       x = "Gender", y = "Proportion",
       fill = "Vote") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q7. Age by vote choice
# ----------------------------------------------------------------------------

ggplot(teds, aes(x = vote, y = age, fill = vote)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  labs(title = "Age Distribution by Vote Choice",
       x = "Vote Choice", y = "Age") +
  theme_minimal(base_size = 18) +
  theme(legend.position = "none")

# ----------------------------------------------------------------------------
# Q8. Correlation matrix
# ----------------------------------------------------------------------------

teds %>%
  dplyr::select(age, edu, income, Taiwanese, Econ_worse, DPP) %>%
  cor(use = "complete.obs") %>%
  as.data.frame() %>%
  rownames_to_column("var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "cor") %>%
  ggplot(aes(x = var1, y = var2, fill = cor)) +
  geom_tile() +
  geom_text(aes(label = round(cor, 2)), size = 6) +
  scale_fill_gradient2(low = "#e74c3c", mid = "white", high = "#2980b9",
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Matrix of Numeric Variables",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# ----------------------------------------------------------------------------
# Q9. Pairs plot
# ----------------------------------------------------------------------------

teds %>%
  dplyr::select(age, edu, income, Taiwanese, Econ_worse, vote) %>%
  ggpairs(aes(color = vote, alpha = 0.5),
          upper = list(continuous = wrap("cor", size = 4)),
          lower = list(continuous = wrap("points", size = 0.5))) +
  scale_color_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  scale_fill_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  theme_minimal(base_size = 14)


# ============================================================================
# PART II: UNSUPERVISED LEARNING
# ============================================================================

# ----------------------------------------------------------------------------
# Prepare data for clustering
# ----------------------------------------------------------------------------

teds_numeric <- teds %>%
  dplyr::select(age, edu, income, Taiwanese, Econ_worse, DPP) %>%
  scale() %>%
  as.data.frame()

head(teds_numeric)

# ----------------------------------------------------------------------------
# Q10. Choosing k: Elbow method
# ----------------------------------------------------------------------------

fviz_nbclust(teds_numeric, kmeans, method = "wss", k.max = 10) +
  labs(title = "Elbow Method: Optimal Number of Clusters") +
  theme_minimal(base_size = 18)

# Silhouette method
fviz_nbclust(teds_numeric, kmeans, method = "silhouette", k.max = 10) +
  labs(title = "Silhouette Method: Optimal Number of Clusters") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q11. K-Means clustering
# ----------------------------------------------------------------------------

set.seed(6323)
km_result <- kmeans(teds_numeric, centers = 3, nstart = 25)

table(km_result$cluster)
round(km_result$centers, 2)

# Visualize clusters
fviz_cluster(km_result, data = teds_numeric,
             geom = "point", 
             ellipse.type = "convex",
             palette = c("#e74c3c", "#2980b9", "#27ae60"),
             ggtheme = theme_minimal(base_size = 18)) +
  labs(title = "K-Means Clustering of TEDS 2016 Voters (k = 3)")

# ----------------------------------------------------------------------------
# Q12. Cluster profiling
# ----------------------------------------------------------------------------

teds$cluster <- as.factor(km_result$cluster)

teds %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    mean_age = round(mean(age), 1),
    mean_edu = round(mean(edu), 1),
    mean_income = round(mean(income), 1),
    pct_Taiwanese = round(mean(Taiwanese) * 100, 1),
    pct_Econ_worse = round(mean(Econ_worse) * 100, 1),
    mean_DPP = round(mean(DPP), 2),
    pct_Tsai = round(mean(vote == "Tsai") * 100, 1)
  )

# ----------------------------------------------------------------------------
# Q13. Clusters vs. vote choice
# ----------------------------------------------------------------------------

ggplot(teds, aes(x = cluster, fill = vote)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("Other" = "#2980b9", "Tsai" = "#27ae60")) +
  labs(title = "Vote Choice Distribution by Cluster",
       x = "Cluster", y = "Proportion",
       fill = "Vote") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q14. Principal Component Analysis
# ----------------------------------------------------------------------------

pca_result <- prcomp(teds_numeric, scale. = TRUE)
summary(pca_result)

# Scree plot
fviz_eig(pca_result, addlabels = TRUE) +
  labs(title = "Scree Plot: Variance Explained by Each PC") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q15. PCA variable contributions
# ----------------------------------------------------------------------------

fviz_pca_var(pca_result, 
             col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE) +
  labs(title = "PCA: Variable Contributions") +
  theme_minimal(base_size = 18)

# PCA individuals by vote
fviz_pca_ind(pca_result,
             geom = "point",
             col.ind = teds$vote,
             palette = c("#2980b9", "#27ae60"),
             addEllipses = TRUE,
             legend.title = "Vote") +
  labs(title = "PCA: Voters Projected onto First Two PCs") +
  theme_minimal(base_size = 18)


# ============================================================================
# PART III: SUPERVISED LEARNING
# ============================================================================

# ----------------------------------------------------------------------------
# Q16. Train-test split
# ----------------------------------------------------------------------------

set.seed(6323)

train_index <- createDataPartition(teds$vote, p = 0.7, list = FALSE)
train_data <- teds[train_index, ] %>% dplyr::select(-cluster)
test_data  <- teds[-train_index, ] %>% dplyr::select(-cluster)

cat("Training set:", nrow(train_data), "observations\n")
cat("Test set:", nrow(test_data), "observations\n")
prop.table(table(train_data$vote))

# ----------------------------------------------------------------------------
# Q17. Logistic regression
# ----------------------------------------------------------------------------

logit_model <- glm(vote ~ age + gender + edu + income + 
                     Taiwanese + Econ_worse + DPP,
                   data = train_data, 
                   family = binomial)

summary(logit_model)

# Evaluation
logit_probs <- predict(logit_model, newdata = test_data, type = "response")
logit_pred <- ifelse(logit_probs > 0.5, "Tsai", "Other")
logit_pred <- factor(logit_pred, levels = c("Other", "Tsai"))

confusionMatrix(logit_pred, test_data$vote)

# ----------------------------------------------------------------------------
# Q18. Decision tree
# ----------------------------------------------------------------------------

tree_model <- rpart(vote ~ age + gender + edu + income + 
                      Taiwanese + Econ_worse + DPP,
                    data = train_data,
                    method = "class")

rpart.plot(tree_model, 
           type = 4, 
           extra = 104,
           main = "Decision Tree: Predicting Vote Choice",
           box.palette = "BuGn",
           cex = 1.2)

# Evaluation
tree_pred <- predict(tree_model, newdata = test_data, type = "class")
confusionMatrix(tree_pred, test_data$vote)

# ----------------------------------------------------------------------------
# Q19. Random forest
# ----------------------------------------------------------------------------

set.seed(6323)

rf_model <- randomForest(vote ~ age + gender + edu + income + 
                           Taiwanese + Econ_worse + DPP,
                         data = train_data,
                         ntree = 500,
                         importance = TRUE)

print(rf_model)

# Evaluation
rf_pred <- predict(rf_model, newdata = test_data)
confusionMatrix(rf_pred, test_data$vote)

# ----------------------------------------------------------------------------
# Q20. Variable importance
# ----------------------------------------------------------------------------

importance_df <- as.data.frame(importance(rf_model)) %>%
  rownames_to_column("variable") %>%
  arrange(desc(MeanDecreaseGini))

ggplot(importance_df, aes(x = reorder(variable, MeanDecreaseGini), 
                           y = MeanDecreaseGini)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Random Forest: Variable Importance",
       subtitle = "Mean Decrease in Gini Index",
       x = NULL, y = "Importance") +
  theme_minimal(base_size = 18)

# ----------------------------------------------------------------------------
# Q21. Model comparison
# ----------------------------------------------------------------------------

results <- tibble(
  Model = c("Logistic Regression", "Decision Tree", "Random Forest"),
  Accuracy = c(
    confusionMatrix(logit_pred, test_data$vote)$overall["Accuracy"],
    confusionMatrix(tree_pred, test_data$vote)$overall["Accuracy"],
    confusionMatrix(rf_pred, test_data$vote)$overall["Accuracy"]
  )
)

results %>% arrange(desc(Accuracy))

# Comparison plot
ggplot(results, aes(x = reorder(Model, Accuracy), y = Accuracy, fill = Model)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = paste0(round(Accuracy * 100, 1), "%")), 
            hjust = -0.1, size = 7) +
  coord_flip() +
  scale_fill_manual(values = c("#e74c3c", "#2980b9", "#27ae60")) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(title = "Model Comparison: Prediction Accuracy",
       x = NULL, y = "Accuracy") +
  theme_minimal(base_size = 18) +
  theme(legend.position = "none")
```

## Multiple Regression Exercise

This assignment applies ideas from ISLR Chapter 3 on linear regression. Multiple regression is used when one dependent variable is modeled using more than one predictor at the same time. In this exercise, the TEDS2016 dataset is used to run a multiple regression model and to create regression plots showing the relationship between the dependent variable and Age, Education, and Income.

## 2. Review ISLR Chapters 3

ISLR Chapter 3 provides the background for this assignment by explaining simple linear regression and multiple linear regression. The chapter shows that regression is used to model the relationship between a response variable and one or more predictors, and that in a multiple regression model each coefficient represents the effect of one predictor while holding the other predictors fixed.

## 3. Load the TEDS2016 dataset

The next step is to load the TEDS2016 dataset into R using the exact code provided in the assignment. This is necessary because the dataset must be available in the R environment before the regression model and regression plots can be created.

```{r, message=FALSE, warning=FALSE}
library(haven)

TEDS_2016 <- read_stata("https://github.com/datageneration/home/blob/master/DataProgramming/data/TEDS_2016.dta?raw=true")
```

## 4. Run the multiple regression model

After loading the dataset, the next step is to run a multiple regression model. Multiple regression is appropriate when the response variable is modeled using several predictors simultaneously. In this case, `PartyID` is used as the dependent variable, while `Age`, `Edu`, and `income` are used as predictors. This model estimates how each predictor is associated with the dependent variable while the other predictors are held constant.

```{r}
lm.fit <- lm(PartyID ~ Age + Edu + income, data = TEDS_2016)
summary(lm.fit)
```

The regression summary provides the estimated coefficients, standard errors, t-values, p-values, and overall model fit statistics. The coefficients show the direction of the relationship for each predictor, the p-values show whether those relationships appear statistically significant, and the residual standard error and multiple R-squared describe how well the model fits the data overall.

## 5. Create the regplot function

The assignment then asks for a function called `regplot` to draw a regression fit plot. This function combines `lm()`, `plot()`, and `abline()` to create a scatterplot and add the fitted regression line in red. Writing the function once makes it possible to reuse the same plotting method for each predictor.

```{r}
regplot=function(x,y){
 fit=lm(y~x)
 plot(x,y)
 abline(fit,col="red")
}
```

## 5a. Run a regplot using Age

The first regression plot uses Age as the explanatory variable and the dependent variable as the response. This makes it possible to visually examine whether the dependent variable tends to increase or decrease as Age changes.

```{r}
regplot(TEDS_2016$Age, TEDS_2016$PartyID)
```

## 5b. Run a regplot using Education

The second regression plot uses Education as the explanatory variable. This plot provides a visual summary of the relationship between Education and the dependent variable and allows it to be compared with the Age plot.

```{r}
regplot(TEDS_2016$Edu, TEDS_2016$PartyID)
```

## 5c. Run a regplot using Income

The third regression plot uses Income as the explanatory variable. This shows whether the dependent variable appears to move upward or downward as Income changes.

```{r}
regplot(TEDS_2016$income, TEDS_2016$PartyID)
```

## 6. What is the problem? Why?

The main problem is that the dependent variable is not a continuous outcome. Instead, it consists of discrete categories, so the points in the regression plots appear in stacked horizontal levels rather than along a continuous scale. Because ordinary linear regression is designed for a quantitative continuous response variable, the fitted regression line is not the most appropriate way to model or predict this type of outcome.

## 7. What can be done to improve prediction of the dependent variable?

Prediction can be improved by using a model that is designed for a categorical dependent variable rather than ordinary linear regression. A better choice would be multinomial logistic regression if the dependent variable has several categories. Another possible improvement would be to recode the dependent variable into fewer meaningful groups before modeling.

**AI Disclaimer:** *I used ChatGPT as a tutor to clarify instructions, troubleshoot errors, and better understand mistakes encountered during the process. It also assisted with wording and formatting; however, the final submission reflects my own work and judgment.*
