---
title: "Assignment 3"
format:
  html:
    theme: darkly
    highlight-style: zenburn
    toc: false
---

<style>

html, body {
  background-color: #060709 !important;
  color: #f2f2ed !important;
}

body {
  margin: 0 !important;
  font-family: "Courier New", Courier, monospace !important;
  overflow-x: hidden !important;
  background-image:
    linear-gradient(rgba(255,255,255,0.025) 1px, transparent 1px),
    linear-gradient(90deg, rgba(255,255,255,0.025) 1px, transparent 1px) !important;
  background-size: 42px 42px !important;
}

.main-container {
  max-width: 1000px !important;
  margin: auto !important;
  padding: 35px 55px 60px 55px !important;
  background-color: transparent !important;
}

.author,
.date,
.quarto-title-meta,
.quarto-title-meta-contents {
  display: none !important;
}

h1.title {
  color: #f2f2ed !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: clamp(2.8rem, 7vw, 6rem) !important;
  font-weight: 800 !important;
  letter-spacing: -5px !important;
  line-height: 0.95 !important;
  text-transform: uppercase !important;
  text-align: left !important;
  margin-bottom: 2rem !important;
}

h1 {
  color: #f2f2ed !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: clamp(2.4rem, 6vw, 4.5rem) !important;
  font-weight: 800 !important;
  letter-spacing: -4px !important;
  line-height: 1 !important;
  text-transform: uppercase !important;
  border-bottom: 1px solid rgba(255,255,255,0.16) !important;
  padding-bottom: 12px !important;
}

h2 {
  margin-top: 4rem !important;
  padding-top: 1rem !important;
  color: #f2f2ed !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: clamp(1.5rem, 4vw, 2.6rem) !important;
  font-weight: 700 !important;
  letter-spacing: -2px !important;
  border-top: 1px solid rgba(255,255,255,0.16) !important;
}

h3 {
  margin: 3rem 0 1rem !important;
  color: #dfff54 !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: 1.25rem !important;
  font-weight: 700 !important;
}

p {
  margin-bottom: 2rem !important;
  color: #9ca5ae !important;
  font-family: Arial, Helvetica, sans-serif !important;
  font-size: 0.95rem !important;
  line-height: 1.7 !important;
}

code {
  color: #dfff54 !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: 0.85rem !important;
  font-weight: 700 !important;
  background-color: transparent !important;
}

pre {
  background-color: #080b0e !important;
  color: #f2f2ed !important;
  border: 1px solid rgba(255,255,255,0.16) !important;
  border-radius: 0 !important;
  padding: 16px !important;
  font-family: "Courier New", Courier, monospace !important;
  font-size: 13px !important;
  line-height: 1.5 !important;
  overflow-x: auto !important;
}

pre code {
  background-color: transparent !important;
  padding: 0 !important;
}

table {
  width: 100% !important;
  color: #f2f2ed !important;
  font-family: "Courier New", Courier, monospace !important;
  background-color: #080b0e !important;
  border-collapse: collapse !important;
}

th {
  color: #dfff54 !important;
  background-color: #0d1115 !important;
  border-bottom: 1px solid rgba(255,255,255,0.20) !important;
  text-align: center !important;
  padding: 10px !important;
}

td {
  color: #f2f2ed !important;
  background-color: #080b0e !important;
  border-bottom: 1px solid rgba(255,255,255,0.10) !important;
  text-align: center !important;
  padding: 10px !important;
}

caption {
  color: #9ca5ae !important;
  font-family: Arial, Helvetica, sans-serif !important;
  padding-bottom: 10px !important;
}

img {
  display: block !important;
  margin: 22px auto 32px auto !important;
  max-width: 95% !important;
}

a {
  color: #dfff54 !important;
}

a:hover {
  color: #ffffff !important;
}

</style>


# 1. Documenting a Chart

## 1.a: Re-running murrell01.R from Assignment 2

```{r}
#| label: murrell01
#| echo: true
#| results: hide
#| fig-show: hold
#| fig-width: 8
#| fig-height: 6

source("murrell01.R")
```


## 1.b: Documenting one chart

I chose the boxplot. The comments explain how each part contributes to the finished chart.

```{r}
#| label: boxplot
#| echo: true
#| fig-width: 8
#| fig-height: 5.5

# Boxplot

# Set margins
par(mar=c(3, 4.1, 2, 0))

# VC boxplots
boxplot(len ~ dose, data=ToothGrowth,
        boxwex=0.25, at=1:3 - 0.2,
        subset=supp == "VC", col="white",
        xlab="",
        ylab="Tooth Length", ylim=c(0,35))

# Add x-axis label
mtext("Vitamin C dose (mg)", side=1, line=2.5, cex=0.8)

# Add OJ boxplots
boxplot(len ~ dose, data=ToothGrowth, add=TRUE,
        boxwex=0.25, at=1:3 + 0.2,
        subset=supp == "OJ")

# Add legend
legend(1.5, 9, c("Ascorbic acid", "Orange juice"),
       fill=c("white", "gray"),
       bty="n")

# Reset margins
par(mar=c(5.1, 4.1, 4.1, 2.1))
```


# 2. Anscombe's Quartet

## 2.a: Re-running anscombe01.R

```{r}
#| label: anscombe-original
#| echo: true
#| results: hide
#| fig-show: hide

# Run anscombe01.R without opening the RStudio data viewer
local({
  View <- function(...) invisible(NULL)
  source("anscombe01.R", local=environment())
})
```

```{r}
#| label: anscombe-data
#| echo: true

## Anscombe (1973) Quartet

data(anscombe)
summary(anscombe)

## Fit all four models

lm1 <- lm(y1 ~ x1, data=anscombe)
lm2 <- lm(y2 ~ x2, data=anscombe)
lm3 <- lm(y3 ~ x3, data=anscombe)
lm4 <- lm(y4 ~ x4, data=anscombe)
```


### Anscombe I

```{r}
#| label: anscombe1
#| echo: true
#| fig-width: 7
#| fig-height: 5

plot(anscombe$x1, anscombe$y1)
abline(coefficients(lm1))
```


### Anscombe II

```{r}
#| label: anscombe2
#| echo: true
#| fig-width: 7
#| fig-height: 5

plot(anscombe$x2, anscombe$y2)
abline(coefficients(lm2))
```


### Anscombe III

```{r}
#| label: anscombe3
#| echo: true
#| fig-width: 7
#| fig-height: 5

plot(anscombe$x3, anscombe$y3)
abline(coefficients(lm3))
```


### Anscombe IV

```{r}
#| label: anscombe4
#| echo: true
#| fig-width: 7
#| fig-height: 5

plot(anscombe$x4, anscombe$y4)
abline(coefficients(lm4))
```


### Four plots together

```{r}
#| label: anscombe-four
#| echo: true
#| fig-width: 9
#| fig-height: 7

par(mfrow=c(2,2))

plot(anscombe$x1, anscombe$y1,
     col="red", pch=21, bg="orange",
     xlim=c(3,19), ylim=c(3,13),
     main="Anscombe I")
abline(lm1, col="blue")

plot(anscombe$x2, anscombe$y2,
     col="red", pch=21, bg="orange",
     xlim=c(3,19), ylim=c(3,13),
     main="Anscombe II")
abline(lm2, col="blue")

plot(anscombe$x3, anscombe$y3,
     col="red", pch=21, bg="orange",
     xlim=c(3,19), ylim=c(3,13),
     main="Anscombe III")
abline(lm3, col="blue")

plot(anscombe$x4, anscombe$y4,
     col="red", pch=21, bg="orange",
     xlim=c(3,19), ylim=c(3,13),
     main="Anscombe IV")
abline(lm4, col="blue")

par(mfrow=c(1,1))
```


## 2.b: Comparing the regression models

```{r}
#| label: regression-table
#| echo: true
#| results: asis

results <- data.frame(
  Model=c("Model 1", "Model 2", "Model 3", "Model 4"),
  Intercept=c(coef(lm1)[1],
              coef(lm2)[1],
              coef(lm3)[1],
              coef(lm4)[1]),
  Slope=c(coef(lm1)[2],
          coef(lm2)[2],
          coef(lm3)[2],
          coef(lm4)[2]),
  R_Squared=c(summary(lm1)$r.squared,
              summary(lm2)$r.squared,
              summary(lm3)$r.squared,
              summary(lm4)$r.squared),
  Residual_SE=c(summary(lm1)$sigma,
                summary(lm2)$sigma,
                summary(lm3)$sigma,
                summary(lm4)$sigma)
)

rownames(results) <- NULL
results[, -1] <- round(results[, -1], 3)

knitr::kable(results,
             format="html",
             caption="Comparison of Anscombe's Four Regression Models",
             align="c")
```

The four regression models have almost identical results. Each model has an intercept around 3.00, a slope around 0.50, an R-squared value around 0.67, and a residual standard error around 1.24. Based only on these statistics, the four datasets look almost the same.


## 2.c: Comparing ways of drawing them

```{r}
#| label: anscombe-different
#| echo: true
#| fig-width: 9
#| fig-height: 7

par(mfrow=c(2,2))

plot(anscombe$x1, anscombe$y1,
     col="deeppink",
     pch=16,
     main="Anscombe I")
abline(lm1, col="navy", lty=1, lwd=2)

plot(anscombe$x2, anscombe$y2,
     col="darkgreen",
     pch=17,
     main="Anscombe II")
abline(lm2, col="purple", lty=2, lwd=2)

plot(anscombe$x3, anscombe$y3,
     col="darkorange",
     pch=15,
     main="Anscombe III")
abline(lm3, col="darkred", lty=3, lwd=2)

plot(anscombe$x4, anscombe$y4,
     col="steelblue",
     pch=18,
     main="Anscombe IV")
abline(lm4, col="gray40", lty=4, lwd=2)

par(mfrow=c(1,1))
```

Although the four datasets have almost identical regression statistics, their plots show very different relationships. Anscombe I has a mostly linear relationship. Anscombe II has a curved pattern. Anscombe III is mostly linear but has an influential outlier. Anscombe IV has most of the observations around the same x-value with one point strongly affecting the regression line. The summary statistics hide these differences in shape, outliers, and influential observations.


# 3. Finetune Without Leaving Base R

## 3.a: Serif font

`par(family="serif")`

## 3.b: Non-default colours

`col="deeppink"` and `col="darkblue"`

## 3.c: My own plotting character

`pch=18`

```{r}
#| label: finetune
#| echo: true
#| fig-width: 8
#| fig-height: 5.5

par(mfrow=c(1,1))

# Serif font
par(family="serif")

# Pressure plot
plot(pressure,
     pch=18,
     col="deeppink",
     main="Pressure versus Temperature",
     xlab="Temperature (Celsius)",
     ylab="Pressure (mm Hg)")

# Add text
text(150, 600,
     "Pressure (mm Hg)\nversus\nTemperature (Celsius)",
     col="darkblue",
     family="serif")
```


# 4. Now Doing It in ggplot2

```{r}
#| label: ggplot2
#| echo: true
#| message: false
#| warning: false
#| fig-width: 8
#| fig-height: 5.5

library(tidyverse)

# Convert pressure to data frame
df <- as.data.frame(pressure)

# Create ggplot
ggplot(df, aes(x=temperature, y=pressure)) +
  geom_point(shape=18,
             color="deeppink",
             size=2.5) +
  annotate("text",
           x=150,
           y=600,
           label="Pressure (mm Hg)\nversus\nTemperature (Celsius)",
           color="darkblue",
           family="serif") +
  labs(title="Pressure versus Temperature",
       x="Temperature (Celsius)",
       y="Pressure (mm Hg)") +
  theme_minimal(base_family="serif")
```


## Base R and ggplot2 Comparison

In base R, I used `par(family="serif")` for the font. In ggplot2, this became `theme_minimal(base_family="serif")`.

The base R plotting character `pch=18` became `shape=18` in `geom_point()`. The base R point colour `col="deeppink"` became `color="deeppink"`.

I did not need to use a `scale_*()` function because the colours and plotting characters were assigned directly instead of being mapped to a variable. If they were mapped to categories, I could use functions such as `scale_color_manual()` or `scale_shape_manual()`.

The base R arguments `main`, `xlab`, and `ylab` became the `title`, `x`, and `y` arguments in `labs()`.

The base R `text()` function became `annotate("text", ...)` in ggplot2.

There is not one exact ggplot2 replacement for `par()`. In ggplot2, those settings are handled through `theme()`, geoms, scales, and other plot settings.


# 5. Pre-hackathon — Team Task

The goal of this task is to replicate the Daily COVID Deaths in Europe chart using the Our World in Data dataset.


## 5.a: Acquire the data

```{r}
#| label: covid-data
#| echo: true
#| message: false
#| warning: false

# Daily COVID counts, Our World in Data

url <- paste0(
  "https://raw.githubusercontent.com/owid/covid-19-data/",
  "master/public/data/owid-covid-data.csv"
)

# Download data
download.file(url, "owid-covid-data.csv")

# Read data
owidall <- read.csv("owid-covid-data.csv")

# Remove OWID aggregate rows
owidall <- owidall[!grepl("^OWID", owidall$iso_code), ]

# Keep Europe only
owideu <- subset(owidall, continent == "Europe")
```


## 5.b: Replicate the chart

```{r}
#| label: covid-chart
#| echo: true
#| fig-width: 10
#| fig-height: 6.5

# Convert date column to Date
owideu$date <- as.Date(owideu$date)

# Keep dates shown in target chart
eu_plot <- subset(
  owideu,
  date >= as.Date("2020-01-01") &
  date <= as.Date("2023-08-31") &
  !is.na(new_deaths)
)

# Set margins and serif font
par(mar=c(7, 6, 2, 2), family="serif")

# Create scatterplot
plot(
  eu_plot$date,
  eu_plot$new_deaths,
  pch=16,
  cex=0.45,
  col="deeppink",
  xaxt="n",
  xlab="",
  ylab="COVID Deaths in Europe (Daily)",
  ylim=c(-200, 6700)
)

# Create dates for x-axis
date_ticks <- seq(
  as.Date("2020-01-01"),
  as.Date("2023-08-01"),
  by="2 months"
)

# Add rotated date labels
axis(
  1,
  at=date_ticks,
  labels=format(date_ticks, "%Y-%m"),
  las=2,
  cex.axis=0.7
)

# Add x-axis title
mtext("Date", side=1, line=5)

# Find high points
label_data <- subset(
  eu_plot,
  new_deaths >= 5000 &
  new_deaths <= 6500
)

# Sort highest to lowest
label_data <- label_data[
  order(label_data$new_deaths, decreasing=TRUE),
]

# Keep one high point per country
label_data <- label_data[
  !duplicated(label_data$location),
]

# Keep highest eight countries
label_data <- head(label_data, 8)

# Label high points by country
text(
  label_data$date,
  label_data$new_deaths,
  labels=label_data$location,
  pos=3,
  cex=0.7
)

# Reset margins
par(mar=c(5.1, 4.1, 4.1, 2.1))
```

The date column is converted into R's `Date` format so that it can be plotted correctly on the x-axis. The data is limited to observations from January 2020 through August 2023, and observations with missing daily death values are removed.

The bottom margin is increased to make room for the rotated date labels. I used a serif font and `deeppink` points to match the target chart. The default x-axis is removed with `xaxt="n"` so that I could create date labels every two months and rotate them vertically.

Finally, I selected the high points in the visible range of the chart, kept the highest observation for each country, and used `text()` to label those observations by country. The labels are placed directly next to the observations instead of being displayed as a legend.


AI Disclaimer: I used ChatGPT as a tutor to clarify instructions, troubleshoot errors, and better understand mistakes encountered during the process. It also assisted with wording and formatting; however, the final submission reflects my own work and judgment.
